\section{Case Studies}

In this section, we discuss some case studies in which we applied
\ourDSL to a variety of data structures.
To do this we collected real-world code
from multiple sources:
we used 25 functions from \hskin{Data.List}
(part of GHC's standard library)
 14 functions from \hskin{Data.Map} (part of the widely used \texttt{containers} package)
and 3 from a \hskin{Union-Find} data structure
implemented as part of an academic Haskell project~\cite{hallahan2019lazy}
following the basic design of~\cite{conchon2007persistent}.

\input{sections/figures/case-study-list-ex.tex}
\mypara{List}
We wrote 25 \ourDSL tests for \texttt{Data.List},
each involving a single function from the library.
The \hskin{zip} test, from the motivating example (Section~\ref{sec:motivating-example}),
is one example of a test for this library.
A second example, for append, \hskin{(++)},
is shown in Figure~\ref{fig:list-append-ex}.
% Here, \hskin{test2} acts as a higher-order function, taking four arguments: 
% (1) the function to test, (2) a symbolic expression for its first argument, 
% (3) a symbolic expression for its second argument, 
% and (4) a symbolic expression for the expected output.  
In the \hskin{(++)} example, the first symbolic example, \hskin;[ex| b*{a} |];,
is the symbolic expression for the 
first input to \hskin{(++)}, generating a random list of integers of length a. 
Similarly, the third argument \hskin;[ex| d*{c} |]; generates a random list 
of integers of length c. Lastly, the symbolic expression \hskin;[ex| b*{a} d*{c} |]; 
stands for the concatenated output of these two lists. 
\ourDSL generates random inputs for input symbolic expressions and 
runs the \hskin{(++)} with those inputs. 
It then compares the output against the output symbolic expression. 

\mypara{Higher-order functions}
\ourDSL currently does not possess a way to generate arbitrary functions, so testing a high order function on arbitrary inputs is impossible. 
However, it is possible to test high order functions through partial application. 
For example, consider the test in Figure~\ref{fig:list-filter-ex}.
The filter function accepts a predicate as its first argument and a list as its second, and returns a list containing only the elements where the predicate is true.
To test this with
\ourDSL,
we apply \hskin{filter}
in a lambda \hskin{\x -> filter (/= x)}.
This fixes filter to removing some specific element,
but allows the symbolic test to vary that element,
and vary the actual list that is passed in.
We can then specify precisely what the output
should be relatively to the input:
we use a labeled union to
select including either an \hskin{x}
or a not-equal-to-\hskin{x} element in the list,
\hskin+(x |{b} (y /= x))*{n}+.
We then can specify
that in the output list, all the \hskin{x}s
should be missing, and all other elements should be present,
\hskin+([] |{b} y)*{n}+.
% This is obviously not a proper high order function since none of the arguments are functions themselves,
% however we can still test the implementation of a high order function, filter in this case, by predetermining what the function arguments should be
% and accepting extra arguments that we can randomly generate to construct them.
 
\input{sections/figures/case-study-lookup-ex.tex}
\input{sections/figures/further-case-studies}

\mypara{Map} \label{subsec:case-studies-map}
To test the effectiveness of \ourDSL with other data structures, 
we also evaluated it on functions from Haskell's \hskin{Data.Map} 
library, which implements \hskin{Map},
a key-value store.
Although \ourDSL's symbolic expressions describe lists,
this is (in many cases) not a barrier to specify tests
for functions of other datatypes,
as conversion to and from lists are common.
\hskin{Map} serves as an example of this strategy.
% We selected Map because symbolic expressions in our 
% approach can generate only lists, and \hskin{Map} provides functionality 
% to convert between maps and lists, making it suitable for our tests.

Figure~\ref{fig:lookup-test} shows a test for the \hskin{Map} function 
\hskin{lookup}. Here, the function 
that we are passing to \hskin{test2} is \hskin{(\x -> lookup x . fromList)} 
of type \hskin{Int -> [(Int, Int)] -> Maybe Int}. The original 
type definition of \hskin{lookup} is \hskin{k -> Map k a -> Maybe a} 
where it takes a key and a map, looks up for the key of some type 
\hskin{k} and returns the value of type \hskin{a} wrapped in a 
\hskin{Maybe}. It returns \hskin{Nothing} if the key does not exist 
in the Map or returns the value by wrapping it in a \hskin{Just} constructor.

Since we only generate lists from symbolic expressions,
we set up the function being tested to convert a generated list to a \hskin{Map} using the 
\hskin{fromList} function. The expression \hskin;[ex| k |]; generates a random key. 
The expression \hskin;[ex| <((. /= k), .)>* <(k,v)> <((. /= k), .)>* |]; 
creates a list of tuples, representing key-value pairs.
\hskin{(. /= k)} enforces that the value generated by the dot is not equal to \hskin{k},
thus we guarantee that the only tuple with $k$ as the first element will be the one generated
by \hskin{<(k,v)>}.
We can then convert this into a \hskin{Map}, using \hskin{fromList},
and then know that the \hskin{Map} associates key \hskin{k} with value \hskin{v}.
Thus, when \hskin{lookup} is called,
it should produce \hskin{Just v},
matching the output symbolic expression.
% So, in the subexpression 
% \hskin{<((. /= k), .)>* } angled brackets perform \hskin{(,)}
% constructor application.
% Within the same sub-expression, \hskin{(. /= k)}
% is actually a predicate that generates a key which is not equal to the one 
% we are looking for.
% Together \hskin{<((. /= k), .)>* }
% generates the list of key-value pair 
% that does not contain key \hskin{k}.
% Sub-expression \hskin{<(k, v)>} 
% generate a tuple \hskin{(k,v)}. 
% Later, the combined result of all the 
% sub-expressions forms the input list. Forming the symbolic 
% expression in the mentioned way, make sure the input 
% list contains only one such key-value pair.

Figure~\ref{fig:additional-map} shows additional
tests for \hskin{Map} functions which compute intersection and difference on keys,
as well as the function \hskin{split k m},
which divides \hskin{m} into a \hskin{Map} of keys smaller than \hskin{k}
and a \hskin{Map} of keys larger than \hskin{k}
(keys that are exactly \hskin{k} are not in either \hskin{Map}.)

\input{sections/figures/case-study-uf.tex}
\mypara{Union-Find}
We further tested \ourDSL on a union-find data structure,
extracted from an existing academic project~\cite{hallahan2019lazy},
and (according to a comment in the code) based on~\cite{conchon2007persistent}.
Union-Find is a data structure for representing a collection of disjoint sets, 
with each set being assigned one element to represent it.
The Union-Find data structure has two primary operations, \hskin{union} and \hskin{find}.
\hskin{union} takes two elements of the structure as parameters, and unions the two sets that they are in.
The find operation takes one element as a parameter, and returns the representative from the set that contains it.
You can determine if two elements are in the same set by comparing the corresponding representatives returned by \hskin{find}.

\par In Figure~\ref{fig:uf-union-test} we see a symbolic test for the union operation.
As with Maps, we
use \hskin{toList} and \hskin{fromList} to convert the data structure we
are testing to lists that can be generated by \ourDSL.
As lists, union-finds are two dimensional. Each element is one of its
non-singleton disjoint sets that it maintains. 
For example, if after initializing a union-find the operations \hskint{union 1 100} and \hskin{union 5 7} were performed, the lists representation would be \hskin{[[1, 100], [5, 7]]}.
The test above shows a function taking two integer inputs that are not equal. 
The third parameter takes a two dimensional list representing a union find. It is an arbitrary set containing \hskin{x} and an arbitrary set containing y, with any number of arbitrary sets between and around them. The output shows \hskin{x} and \hskin{y} in the same set, having been unioned together.

% \par Since the output contains many dots, we call this a partially specific test, in that while certain properties of the output are known, namely \hskin{x} and \hskin{y} being in the same set, the exact value of the output is not entirely clear.

\mypara{Threats to Validity}
\textit{Internal Validity}
A limitation of \ourDSL is its inability to reference label's outside of its current evaluation context.
In other words, when working with \hskin|x*{y}|, 
\ourDSL only has access to one value for $x$, determined by the value of $y$.
This does not stop us from writing many useful tests
for a variety of data structures, but does limit our ability to
test certain functions (\hskin{sort}, for example.)
While we considered three data structures, and were able to write useful tests for some functions in all of them,
it is possible the library could be less applicable to other data structures.
% This makes it impossible to consider permutations of \hskin|x*{y}|, find the minimum $x$ generated, or determine if there are duplicate $x$ values.
% A problem when writing tests for certain functions in \hskin{Data.Map}
% was that we couldn't ensure uniqueness of \textit{all} keys being inserted or account for
% the fact that converting the map to the list might change the order of the tuples around.

\textit{External Validity}
\ourDSL benefits from several features of Haskell:
GADTs, type level functions, and QuasiQuotations.
% that allow us to
% to do preprocessing during compilation to help catch errors faster.
While these features are not universally accessible in other languages,
we believe the overall approach of \ourDSL could still be adapted.
For example, in a 
dynamically-typed language like Python, it would be impossible to tell if a symbolic expression is well formed before runtime.
At the same time, it seems unnecessary to check anything before runtime, since dynamic typings overall philosophy is to keep running until
something forces it to crash.
This means that \ourDSL could be implemented in a dynamically types language
with all
compile time errors shifted runtime errors instead.
We could achieve similar results in a (different) statically typed languages
via type casting and proper error handling with try-catch blocks.