\section{Related Work}

\mypara{Property-Based Testing}
In this work, we consider a new approach to property-based testing
based on writing function specifications in a \textit{symbolic expressions} eDSL.
We explore this direction in the functional programming world,
where in particular QuickCheck~\cite{claessen2000quickcheck},
alongside other approaches such as SmallCheck~\cite{runciman2008smallcheck}
and Target~\cite{seidel2015type},
has led to property-based testing becoming a popular paradigm.

QuickCheck~\cite{claessen2000quickcheck} is often credited as popularizing
property-based testing and inspiring
property-based testing libraries~\cite{maciver2019hypothesis,benac2014jsongen,denes2014quickchick,seidel2015type}.
However,
the earliest work on property-based testing~\cite{fink1994property,fink1997property}
was actually on C code in the Unix community,
where property based testing was used to check security properties and search for exploits.
It introduced a specialized language, TASPEC, for writing testing oracles and specifications,
with an eye specifically toward security.

Despite this, the statement that QuickCheck~\cite{claessen2000quickcheck},
a Haskell property-based testing library,
was highly influential on much, if not all,
of the work that follows is quite fair.
As shown earlier in the paper \bill{Where?}
QuickCheck allows writing specifications over functions using mostly ordinary Haskell code,
using typeclass instances to specify how to generate random input values.
\bill{More on QuickCheck}
SmallCheck~\cite{runciman2008smallcheck} and Target~\cite{seidel2015type}
are alternative Haskell property-based testing library.
The design of SmallCheck resembles the design of QuickCheck,
except that, instead of being tested on randomly chosen inputs,
functions are tested on inputs of increasing size.
Target tests specifications written as \textit{refinement types}
i.e. regular types annotated with decidbale predicates.
This refinement types can also be verified with the LiquidHaskell~\cite{vazou2014refinement} verifier;
a major motivation of Target is to act as a complementary tool to LiquidHaskell;
for example, by allowing mixing of verification on ``easy to verify'' code
and testing on ``hard to verify'' code,
and by finding counterexample when verification is failing because of an \textit{actual} bug.

Adaptions of the core ideas of QuickCheck
exist for a wide variety of other languages, both functional and imperative;
as a (much abbreviated) list,
we note QuickCheck for Isabelle~\cite{bulwahn2012new},
QuickCh\textit{i}ck for Rocq~\cite{denes2014quickchick},
QuickCheck for Rust~\cite{derbenwick2025property},
Hypothesis for Python~\cite{maciver2019hypothesis},
PyModel for Python~\cite{jacky2011pymodel},
and 
Jsongen for json web services~\cite{benac2014jsongen}

% \mypara{White box and black box Input Selection}
% As an alternative to property-based testing,
% functions may be tested by selecting fixed input/output pairs to check in a test suite.
% Methods for selecting such pairs have been studied,
% and can broadly be split into \textit{white box} and \textit{black box} methods~
% \cite{nidhra2012black}.

\mypara{Symbolic Expressions}
Symbolic expressions can be viewed as mini-programs
that generate (related) function inputs and outputs.
Alternatively, they can be seens as abstractions or generalizations
of the inputs themselves.
This second view,
of the expressions as abstractions,
suggests a relation to the manipulation of symbolic program values performed
by program analysis tools.
The most direct link to our work
in this domain is synthesis with abstract examples~\cite{drachsler2017synthesis}.
The ``abstract examples'' in question
represent
strings using a regular expression like language,
extended with labeled kleene stars
and variables.
Unlike \ourDSL, these examples are used purely as an \textit{output mechanism}--
the synthesizer asks the user if certain abstract example input/output pairs
correspond to the function being synthesized, and the user answers yes or no.
The abstract example language in~\cite{drachsler2017synthesis}
is also significantly less sophisticated than ours;
kleene stars are only allowed to repeat a single character,
variables correspond to a single character,
and there are no further features
(i.e. no labeled unions or nested lists.)

Other uses of symbolic representations are further removed,
but still worth considering.
In the world of functional programming,
symbolic execution tools such as
Rosette~\cite{torlak2014lightweight}, G2~\cite{hallahan2019lazy}, SmtEn~\cite{uhler2014smten}, and CutEr~\cite{giantsios2015concolic}
exploit such reasoning to support finding program bugs and (in some cases) verify programs.
These tools general reserve symbolic representations for \textit{internal} use,
communicating information to the user by generating corresponding concrete values.
