% \subsection{GADTs and Type Families}

\mypara{GADTs and Type Families}
A \textit{generalized ADT} (GADT)
is a parametric ADT which allows the \textit{constructor}
to select instantiation of type arguments.
One classic example is a length-indexed \hskin{Vec}:
\begin{smallhskcode}
data Nat = Z | S Nat
data Vec (n :: Nat) a where
    Nil :: Vec Z a
    (:>) :: a -> Vec n a -> Vec (S n) a
\end{smallhskcode}
This code declares an ADT, \hskin{Nat}, to represent the natural numbers,
and a GADT, \hskin{Vec}, which is length-indexed.
This means that the \textit{length} of the vector is encoded into it's type:
in the declaration of \hskin{Nil}, we say the \hskin{Vec} has length \hskin{Zero},
in the declaration of \hskin{:>}, we take a tail of any length \hskin{n},
and then return a \hskin{Vec} with length \hskin{Succ n}.

A type family is a type-level function in Haskell.
During compilation, all instances of the type family are translated into concrete types according to the function.
This is often helpful alongside GADTs.
An append function for \hskin{Vec}'s would need to have the type:
\begin{smallhskcode}
    (+++) :: Vec m a -> Vec n a -> Vec (m + n) a
\end{smallhskcode}
Where \hskin{+} is a type family to add the two list lengths.

In the following, we make use of the \hskin{Nat} type.
To add readability, we use roman numerals,
rather than the \hskin{Nat} constructors.
% \evan{maybe talk more about why this is important? probably don't need an example since we're about to give some anyways.}