% \section{Overview}
\input{sections/figures/host-language}
\input{sections/figures/host-test-rule}

Our approach to property-based testing
uses an embedded domain specific language (eDSL), \ourDSL,
within a \textit{host language}-- 
a traditional programming language that a user wishes to test code in
(our practical implementation uses Haskell as its host language, for example.)
The users write eDSL code in what we refer to as the \textit{surface language}--
this language is designed to be expressive and ergonomic to program in.
The surface language is then compiled to a smaller \textit{core language},
which is then used to actually generate tests.
We use \expr to denote generic host language expressions,
\surface to denote generic surface language expressions,
and \core to denote generic core language expressions.


Figure~\ref{fig:the-host-language}
shows a small snippet of the essential elements in the host language.
The most straightforward of these is
standard \textit{cons-lists},
where we write \ccons for cons and \nil for the empty list.
We assume the language additionally contains
other (possibly user defined) algebraic data types,
each with a corresponding set of data constructors
which may be applied to some number of arguments,
$D\usp\vec{\expr}$.
Our formalism does not specifically handle primitive types, such as ints
(although our implementation does.)
For the purpose of the formalism, 
one can consider primitives
as a special case of
an ADT,
with one constructor
(each with no arguments)
for each possible value.

DSL compilation, $\compileDSL{\surface}$,
and tests, $f\usp\vec{\core} \testEq \core$,
represent functions we implement in the host language to interface with the eDSL.
DSL compilation, $\compileDSL{\surface}$ translates the surface language expression
$\surface$ into a corresponding core language expression $\core$.
(This translation is described in Section~\ref{subsec:the-surface-language}.)
A test $f\usp\vec{\core_1} \testEq \core_2$,
takes a host language function $f$,
a list of (core) symbolic expressions $\vec{\core_1}$ corresponding to
the function's arguments,
and a core symbolic expression $\core_2$ corresponding to the function's output.
The test indicates that $f$ should be run on random host language inputs,
\textit{generated} according to $\vec{\core_1}$,
and then the returned value should be \textit{validated}, or checked for correctness,
according to the output $\core_2$.
\ruletest, in Figure~\ref{fig:test-call-rule},
formalizes this process,
using relations that we will formally define in Sections~\ref{subsec:generation-rules} and~\ref{subsec:validation-rules}.
For now,
ignore the grayed out sections of the rule.
\gens is used to \textit{randomly generate}
host language values that correspond to a symbolic expression. 
Thus $c_1 \gens e_1$ to $c_k \gens e_k$
says that
host language values are generated for each argument to the function being tested.
The function is then applied to these arguments,
yielding a host language expression $e_\mathit{out}$.
\gennedby is used to validate that
a host language list matches a symbolic expression.
In particular,
$e_{\mathit{out}} \gennedby \core_{\mathit{out}}$
returns a host language list $xs$,
we write this as
$e_{\mathit{out}} \gennedby \core_{\mathit{out}} \gennedarrow xs$.
If this list is empty,
the test passed (\testpasses),
if the list is not empty then the test failed (\testfails),
and
$xs$ corresponds to some suffix of $e_{\mathit{out}}$ that could not be validated.

We now consider generation and validation in depth.
Sections~\ref{subsec:the-core-language} and~\ref{sec:mappings}
introduce the core language over which generation and validation work,
and introduce key parameters used during generation and validation.
Section~\ref{subsec:generation-rules}
describes how we generate random values,
and Section~\ref{subsec:validation-rules}
describes how we validate output values.
Finally, Section~\ref{subsec:the-surface-language}
covers how we convert the surface language into the core language.