\subsection{The Core Language} \label{subsec:the-core-language}
\input{sections/figures/core-language}

Figure~\ref{fig:the-core-language}
shows the \textit{core language},
a minimal language for symbolic expressions.
Given what we have introduced so far,
much of the language is likely unsurprising--
we have labeled unions and kleene closures,
assignment to introduce labels,
and labels themselves.
Constructor applications
could simply be literals,
or could be other constructors from other types in the host language--
we in particular consider algebraic datatypes.
\bill{Maybe nail down a bit more.}

The core language has a number of simplifications relative to the surface language.
First, the core language has no
unlabeled unions or Kleene closures.
Second,
all bindings must have a constraint
(although trivially, the constraint may just be $\mathit{True}$.)
Third, the core language is more explicit about when
elements are being cons'ed onto a list (via \cons),
or when two lists are being concatenated together (via \concat.)
Finally, we require that all labels are explicitly introduced by a binding
(in \textit{some} symbolic expression--
this might be an expression for a previous argument,
for example)
before being used.
In Section~\bill{???},
we will show how we translate the (more flexible) surface language
to satisfy the rules of the (more restrictive) core language.

\mypara{Scoping}
In considering scoping, we must keep in mind that we typically have
\textit{multiple} symbolic expressions which \textit{share}
the same symbolic variables.
For example, the symbolic expressions for an argument and return value
(in particular, this describes the identity function on lists):
\[
\textbf{argument: }(k \Longleftarrow \mathit{True}) \asgnarrow ((x \Longleftarrow \mathit{True}) \asgnarrow x)*_k
\quad
\textbf{return: }x*_k
\]
are valid, because $k$ and $x$ are both introduced in the first expression,
and therefore (the same) $k$ and $x$ can be used in the second expression.
Making this work requires some slightly unusual--
but we think still intuitive--
scoping rules. In particular:
\begin{enumerate}
\item We require all variable names to be syntactically unique.
This avoids any confusion from i.e. two $x$'s being bound in one expression,
and then an $x$ (ambiguously) appear within a later expression.
\item \label{item:kleene-union} If $x$ is bound in a expression $c$
and used in a different expression $c'$,
the \textit{Kleene-union context} (explained below) of the binding of $x$ in $c$
must be a (non-strict) subset of the Kleene-union context of the use of $x$ in $c'$.
\item Within an expression that binds $x$,
$x$ may be used anywhere to the right of the binding,
as long as point~\ref{item:kleene-union} is obeyed.
\end{enumerate}

Point~\ref{item:kleene-union} clearly requires more explanation.
The \textit{Kleene-union context} is the set of all labels
on Kleene closures and unions that syntactically surround
a particular point in a program.
We write $\kub{c}{\lab}$
to refer to the Kleene-union context of the binding of a label $\lab$ in in core language expression $c$,
and write $\kuv{c}{\lab}$ to refer to the union of the  Kleene-union contexts of the uses of \lab as a variable in $c$.
For example, if we say that $c$ is (to simplify the expression, we assume that $a$ and $b$ are in scope)
\[
((x \Longleftarrow \mathit{True}) \asgnarrow x) |_a ((y \Longleftarrow \mathit{True}) \asgnarrow y*_b)
\],
then $\kub{c}{x} = \{ a \}$
and  $\kub{c}{y} = \{ a, b \}$.
In point~\ref{item:kleene-union},
requiring the Kleene-union context of $x$'s (resp. $y$) binding in $c$ to be a subset of
the Kleene-union context of $x$'s (resp. $y$) useage in (some) $c'$
ensures that the labels will actually have a value to refer to.
For example, with the above $c$, if $c'$ is
\[
((x x) |_a (y y)*_b) |_z []
\]
This would be allowed (assuming $z$ is in scope),
as we have $\kub{c}{x} = \{a\} \subseteq \{a, z\} = \kuv{c'}{x}$
and we have $\kub{c}{y} = \{a, b\} \subseteq \{a, b, z\} = \kuv{c'}{y}$.
On the other hand, if $c'$ was
\[
((x y) |_a (y y)*_b) |_z []
\]
this would not be allowed,
as $\kub{c, y} = \{a, b\} \not\subseteq \{a, z\} = \kuv{c'}{y}$

\input{sections/figures/type-language}
\mypara{Typing}
Figure~\ref{fig:the-type-language-grammar}
shows a small language for types \type
We distinguish between \textit{base types}, \base,
which represent arbitrary types from the host language,
and \textit{list types}, $[\tau]$.
Figure~\ref{fig:the-type-language-rules}
shows a subset of the typing rules.
The core language really desribes the ability to generate (or match to) lists,
as such,
the typing rules are really giving the type of a list generatable from the expression.
\tyunion
states that a union type must choose between two values of the same type.
\tystar says that a Kleene closure must be applied
to a list type $[\tau]$,
and returns another list of type $[\tau]$
(which works by concatenating all lists generated by repetition together.)
\tybind says that $(\lab \Longleftarrow b) \asgnarrow \core$ has type \type
if $\core$ has type \type.
This might seem surprising if one thinks of binding as if it was a lambda--
note that (as implied by the typing rule),
it is a statement that \textit{some} arbitrary value
should be bound to \lab should be introduced
and used in $\core$.

% \par While the user works with the surface language, the core language (see figure~\ref{fig:the-core-language}). 
% The core language is depth-indexed, meaning that the depth of the core AST is part of its type. 
% The depth-index indicates the nested depth of the list it will generate. 
% A depth of 0 indicates that the result will be a singleton not a list. 
% We use this feature to ensure that all expressions are well formed before runtime. 
% Furthermore, we can put constraints on what depths are allowed for each constructor. 
% For example, the $\concat$ operator is for concatenation which requires 2 lists. 
% So we can ensure that its two arguments have the same depth, and that it is non zero.

