\subsection{Generation Rules} \label{subsec:generation-rules}
\input{sections/figures/generation-rules}
\par We now consider generating random input values from symbolic expressions.
To do this, we define
a relation
$\inp \lab \matches x, \zout$.
$\pathvar$ is the \textit{path} and $\env$ is the \textit{environment};
we elaborate on each of these in the following.

Understanding the path requires noting that,
in generating lists (or, when we come to it later in Section~\bill{whatever}, validating lists)
we may repeatedly use an expression via a kleene closure, i.e. given $r*_\lab$,
the expression $r$ may be repeated an arbitrary number of times.
The path is a set of label-to-int mappings that tracks this repetition--
on first entering a kleene closure $r*_\lab$,
$\lab \rightarrow 1$ is entered into the path.
On each successive $n^{th}$ iteration, this mapping is updated to $n$.

The environment tracks how generation is assigning labels to values via bindings,
$(lab \Longleftarrow p(args, \lab)) \mapsto e$
however note that the same label may be re-assigned different values in each iteration of a
(possibly nested) kleene closure,
thus, the environment is really a mapping of (label, path) pairs to values.



\begin{example}
Consider the expression:
\[
(n \Longleftarrow \mathit{True}) \mapsto ((x \Longleftarrow \mathit{True}) \mapsto x)*_n
\]
One list that could be generated from this would be:
\[
[8, 4, 17]
\]
The corresponding environment would be:
\begin{align*}
\{ &(n, \{\}) \rightarrow 3, \\
&(x, \{ n \rightarrow 1 \}) \rightarrow 8, \\
&(x, \{ n \rightarrow 2 \}) \rightarrow 4, \\
&(x, \{ n \rightarrow 3 \}) \rightarrow 17 \\ \}
\end{align*}
$n$ occurs outside of all kleene closures, and thus has an empty context.
$x$ is chosen to be $8$, $4$, and $17$, respectively,
on the $1^{st}$, $2^{nd}$ and $3^{rd}$ iterations of the unrolling of the kleene closure indexed by $n$,
and thus has a separate value for each of these three contexts.
\end{example}
% Since a single variable may take on multiple values in different contexts,
% the environment really maps labels to dictionaries which map contexts to values.

Figure~\ref{fig:generation-rules}
shows the generation rules.

\mypara{Lists Constructors}
\RuleEmpty, \RuleCons, and \RuleConcat
structurally match on a symbolic list,
recursively generating a concrete list's subcomponents,
and then appropriately composing those subcomponents together.
\RuleEmpty is the simplest of these,
simply saying that an empty list
can only generate an empty list
($\inp \nil \matches \nil, \zout$.)
\RuleCons generates a single element
for the head of the list
and generates the tail of the list,
and composes them with \cons.
Finally, \RuleConcat generates
two separate lists,
and concatenates them together with \concat.
\bill{Evan, it looks like the recursive
applications of $\matches$ on the top
of \RuleConcat and \RuleCons
are a bit malformed--
can you please look at those?
Also, we need to explain in this paragraph
why the context is ``threaded through'' the recursive calls.
(Is the environment threaded through? I think we don't need to have it threaded?
But explain that also.)}

\mypara{Literals}
\RuleLiteral matches on a specific literal--
in a similar fashion to \RuleEmpty,
this rule simply straightforwardly says that a literal 
can only generate exactly that literal.

\mypara{Assignment and Variables}
\RuleAssign
handles binding
of a new label,
$(\lab \Longleftarrow p(args, \lab)) \mapsto e$.
To do this,
it repeatedly generates arbitrary values of the appropriate type
\bill{how do we know appropriate type?},
until it finds a value that satisfies the constraint $p(args, \lab)$.
Once such a value is found,
the label is bound and returned in a new environment,
$\outw{\lab}{x}$.
Given a label appearing in the expression,
\RuleVariable then allows looking up that label in the environment,
and generating the mapped-to value.

\mypara{Union}
\RuleUnionLeft and \RuleUnionRight
handle the $\|_\lab$ operator.
Note that which rule applies
is based on the (previously randomly chosen)
value of $\lab$,
hence getting us an (indirectly) non-deterministic choice.

\mypara{Kleene Closure}
\RuleStar handles kleene closures, $r*_\lab$.
To do this, it looks up $\envget{\env}{\lab} = n$ in the environment,
and then recursive applies \matches to generate $n$ lists that match $r$,
i.e.
$\forall 1 \leq i \leq n . (\env_{i - 1}, \envput{\pathvar}{\lab}{i}) r \matches xs_i, \env_i$.
Note that (1) the environment is threaded through all these calls,
to ensure that we record all variable bindings performed within each regular expression
and (2) as covered in our description of the path above,
this is the one place in the rules where we \textit{modify} the path,
by, on each iteration, setting the label \lab to point to the current iteration count $i$.

\mypara{Apply}
\bill{Evan, the rule might need to be adjusted a bit? Maybe you could also take a stab at writing something here?}
