\subsection{Validation Rules} \label{subsec:validation-rules}
\input{sections/figures/validation-rules}

Once we have a run a function on randomly generated inputs,
we must \textit{validate} that the function's output
matches the output symbolic expression.
We define a relation
$\valinp xs \gennedby e \gennedarrow ys, \valout$,
such that $xs$ is a concrete list and $e$ is a symbolic expression
that we try to match to a \textit{prefix} of $xs$.
(We will discuss the additional $\predtrack$ parameter further on in this section.)
$ys$ is then the \textit{suffix} of $xs$ that was not matched to $e$.
Another way of looking at this
is that
$\valinp xs \gennedby e \gennedarrow ys, \valout$
says that there is some $xs'$ such that $xs = xs' :++ ys$,
and $xs'$ matches the symbolic expression $e$.
This allows ``chaining'' together validation steps.
For example, consider checking if some concrete list $xs$ matches a symbolic expression
$(r_1 \concat r_2) \concat r_3$.
We would first check if, for some $ys$
$\valinp xs \gennedby r_1 \concat r_2 \gennedarrow ys, \valout$.
We would then pass $ys$ to $r_3$,
$(\env', \pathvar, \predtrack') ys \gennedby r_3 \gennedarrow zs,( \env'', \predtrack'')$.
Finally, we would then know the match on $(r_1 \concat r_2) \concat r_3$ succeeded as a whole
if and only if $zs$ is the empty list--
\textit{every value} of $xs$ was matched to the symbolic expression,
with no suffix left over.


Figure~\ref{fig:validation-rules} formalizes the validation rules.
Note that the actual operation of the validation rules is quite different from that of the generation rules.
During generation,
we randomly generate values at each binding and make use of those values later in constructing
a list.
During validation, we instead search for \textit{any} instantiation of bound variables
that will allow the match to succeed.
This requires delaying instantiation of variables until they are \textit{used},
and can be determined to correspond to some concrete expression.
In some cases, we must explore multiple possible matches
(for example, both branches of a union.)
This in reflected in the fact that the rules are \textit{not} entirely syntax-directed--
in the case that multiple rules apply,
the validator tries applying all of them,
in sequence,
until either (1) it finds a rule application that allows the match to succeed
or, (2) all possible rules, and thus the test as a whole, fail.

\mypara{Constraint Store}
Because we are deferring binding of variables until the variable is used,
we require a different handling of constraints.
This is the purpose of the additional \predtrack parameter,
which we call the \textit{constraint store}.
The constraint store is used to record constraints at assignments,
and then look up and check those constraints when actually using a variable.
For the same reasons as the environment,
the constraint store works as a map on variable names and the path
(filtered by the dependency map.)
We use identical syntax for lookups and insertions to the syntax we use for the environment.

Unfortunately, it is not possible to use \gennedby to get the values of the arguments
that must be passed to the predicate,
as there are no corresponding expressions in the host language list being validated.
To get around this when validating, we impose a syntactic condition on constraint arguments that
we do not impose during generation:
a constraint argument may only contain variables that were bound
during a previous generation step.
This allows us to \textit{deterministically} evaluate
arguments via \gens,
as the only non-deterministic rule for \gens is \RuleAssign.
% besides this, we construct the rules such that a list will be validated against a symbolic expression
% if and only if the list could be generated from the symbolic expression.

We now discuss the rules in Figure~\ref{fig:validation-rules} in detail.

\mypara{List Constructors}
\ValidateCons, \ValidateConcat, and \ValidateEmpty
structurally match on a symbolic expression and a concrete list.
\ValidateCons
works very similarly to the corresponding generation rules,
\RuleCons, requiring the head and tail of the concrete list to match
against a head and tail in the symbolic expression.
Note that we require the head to \textit{completely} match
(the leftover suffix must be the empty list),
while the tail may pass on some elements to following pieces of the symbolic expression.
\ValidateEmpty allows an empty symbolic expression to match on any list,
returning the entire list as a suffix.
\ValidateConcat ``non-deterministically'' splits the list $xs$
to match to each of the two sub-expressions in $r_1 \concat r_2$,
passing the suffix obtained from $r_1$ to the expression for $r_2$.
In practice, this non-determinism corresponds to an implementation sequentially (or in parallel)
trying every possible point to split.
While this might seem potentially expensive,
in practice most split points can be quickly discarded based on matching failures in the sub-expressions.

% \mypara{Literals}
% \ValidateLiteral matches a literal $c$ in the list to that same literal $c$ in the symbolic expression.
% As with \ValidateCons and \ValidateEmpty,
% this is not notably different then the corresponding generation rule,
% \RuleLiteral

\mypara{Assignment and Variables}
As mentioned above,
\ValidateAssignment
takes a binding
$(\lab \Longleftarrow p(args, \lab)) \mapsto e$
but does not really do an assignment,
as this will be handled when we use the variable \lab.
This is because we want to put off assigning variables
in order to better align them to the concrete expression
(we will see how this works shortly when discussing \ValidateVarTwo.)
Also as discussed above though,
we apply generation \gens to all the arguments $args$,
obtaining values $args'$
and we store $p(args')$ in the constraint store.
This will allow us to check that,
when we \textit{do} assign \lab a value,
that value satisfies the constraint.

\ValidateVarOne and \ValidateVarTwo
handle matching labels, $v \gennedby \lab$.
\ValidateVarOne applies if the label being matched
is already in the environment--
in this case,
the rule checks that $\envget{\label}$ is indeed $v$.
\ValidateVarTwo allows a label $\lab$ that is not the environment
to match a value $x$.
To do this, it looks up \lab along the current path in the constraint store,
$p = \predtrack_\pathvar(\lab)$ and applies this check to the possible assignment to lab,
$p(x)$.
If the constraint is satisfied,
the value is inserted in the environment to ensure that
mapping of the label is consistent
(i.e. if the variable appears elsewhere in the expression, then \ValidateVarOne will be applied.)

\mypara{Union}
We have four rules to handle a union $r_1 |_\lab r_2$.
Two of these rules,
\ValidateUnionSolveLeft and \ValidateUnionSolveRight,
handle the case where $\lab$ is not bound in the environment.
\ValidateUnionSolveLeft tries to match
the list against the left side of the union $r_1$,
binding $(\pathvar, \lab)$ to $0$ in the environment.
Correspondingly,
\ValidateUnionSolveRight tries to match against $r_2$,
binding $(\pathvar, \lab)$ to $1$.
Both rules use the constraint store, just as is done in \ValidateVarTwo.
\ValidateUnionLeft and \ValidateUnionRight
handle the case where $\lab$ is bound in
the environment.
In this case, no branching will occur,
as $\envget{\lab}$
will determine which of the two rules is applicable. 

\mypara{Kleene Closure}
Much like for the union,
we have seperate rules for kleene closures $r*_\lab$
depending on whether \lab is bound in the environment.
\ValidateStarSolve handles the case where
the label is not bound.
Given a concrete list $xs$
and the symbolic expression $r*_\lab$,
the rule
allows matching $r$ against $x$
some arbitrary number $n$ times,
i.e. setting $\env_0 = \env$,
the match is computed as
$\forall 1 \leq i < n . (\env_{i}, \pathput{\lab}{i}) ys_i \gennedby r \gennedarrow ys_{i + 1}, \env_{i + 1}$.
Again, \ValidateStarSolve use the constraint store to ensure the value selected for $n$ satisfies any stored constraint.
This roughly corresponds to a loop in which
(1) $xs$ is matched to $r$, obtaining some unmatched suffix $ys$,
and then (2) $xs$ is assigned the value $ys$, allowing another match to occur.
The system branches to explore each possible value of
$n$.
\ValidateStar handles matches on stars $*_\lab$ when \lab is mapped in the environment.
This works essentially the same way as in \ValidateStarSolve,
except that no branching is needed, as the value of $n$ is known.

\mypara{Apply}
\ValidateApplication
validates data constructor applications.
The rule works by checking that each
concrete argument matches each symbolic expression \textit{completely};
that is, we require that there be no left over suffix,
as there are no sub-expressions following the arguments to pass this suffix onto.
(Note that a use of the \ValidateApplication rule is typically
going to occur as either the root of a tree,
or a subtree following directly from matching a list head in \ValidateCons.)

% \bill{Somewhere note that predicates not usable in output}

% \par Validation happens slightly different than generation.
% Firstly, there are no constraints on variables during the validation phase.
% The purpose of validation is, given a symbolic expression and a list, to determine if
% it is possible (following the rules of generation) that the symbolic expression could
% generate the list (i.e $\forall e, xs. e \gens xs \Longleftrightarrow xs \gennedby e$).
% When we come across a label that has no value, we must try every value that label could
% take and check if one of them lets the output match the expression. If it can, we say the
% list was validated by the expression.



