\subsection{Evaluation Parameters}
\label{sec:mappings}
Before getting to the actual rules for generation and validation,
we must introduce three key parameters used in these rules.
These are the \textit{dependency map},
the \textit{path},
and the \textit{environment.}
These three constructs work together
to allow us to track values of labels in symbolic expressions.
As also discussed when considering scope,
the value of a variable may depend on the value of surrounding variables.
In particular, this happens when a variable is assigned inside a Kleene closure.
For example the expression $a \asgnarrow (b \asgnarrow b)*_{a}$ will match any list of elements. 
% The elements won't necessarily all be equal even though they were all matched against $b$,
% because $b$ is bound within the Kleene closure,
% and is allowed to take on different values depending on which iteration $a$ we are on.
% Thus the value of $b$ \textit{depends on} $a$.
If we use $a$ as a label on multiple Kleene closures,
and $b$ within each closure,
the values for $b$ all correspond on each iteration.
Thus, at runtime, we must track and store what value for $b$ was chosen for each value of $a$.
The \textit{dependency map} and \textit{path}
help us do this,
and the \textit{environment} is where we actually store the runtime values.

% This is why we use the context so that $b$ can take on a different value for each iteration of $a$.

\mypara{Dependency Map} 
% In a symbolic expression we sometimes see variables defined in the context of other variables.
% This primarily happens when a variable is declared inside a Kleene closure. 
% However, if $b$ is referenced without $a$ being in context, it becomes ambiguous which $b$ should be used.
To track dependencies between variables, we introduce the \textit{dependency map} $\deps$. 
The dependency map is a mapping that keeps track of which variable's runtime values
can vary based on Kleene closure labels.
In the example above, $a$ doesn't depend on anything, and $b$ depends on $a$. 
This means that it is illegal to reference $b$ unless $a$ is in the current context--
such illegal accesses are rejected with an error.
Determining the dependency map requires computing
the \textit{Kleene context}--
analagous to the Kleene-union context in 
Section~\ref{subsec:the-core-language},
except considering only labels on Kleene closures, and \textit{not} labels on unions.
Note that, as unions do not cause
repetition,
labels bound within a union (but not within a closure,)
will only take on a single value.
This is why the dependency map ignores the union labels (unlike the scoping rules.)

% \textit{Dependency Map Computation}
% The dependency map is computed for a group of symbolic expressions that represent a function's
% inputs and outputs \textit{once},
% before we begin generating or checking values. 
% The computation traverses the first expression keeping track of a \textit{context} \depscontext
% of all Kleene star labels that it has passed through. 
% When it comes across an assignment for some variable $\lab$, it adds an entry
% pointing from the label to the current context $\deps[\lab \rightarrow \depscontext]$.
% Once it finishes traversing the first tree, it has a dependency map for the first tree but not the whole function. So it
% threads the result from the first traversal and repeats the process with the subsequent expressions.
% \bill{Rules???}

% A variable may be used
% in multiple contexts throughout all the expressions, but the first use of the variable is the one that ends up in the final
% dependency map.

% \textit{Dependency Map Use}
% The dependency map is used for environment lookups and updates. 
% First, recall that every key in the environment is a context-label pair that gets mapped to a value.
% So whenever a lookup or update occurs, the first thing that happens is we verify that all the label's
% depencies are satisfied by the context that was passed in. Then we filter out all variables from the
% context that are not in the label's dependencies. This gives us $\pathvar'$ which contains only the
% relevant context for the label. This relevant context is what is actually used to lookup values and
% insert pairs into the environment. 
% \evan{Should I give an example here to demonstrate why these steps are necessary?}

\mypara{Path}
% $\pathvar$ is the \textit{path} and $\env$ is the \textit{environment}; we elaborate on each of these in the following. 
% Understanding the path requires noting that, in generating lists (or, when we come to it later in Section~\bill{whatever}, validating lists) 
% we may repeatedly use an expression via a kleene closure, i.e. given $r*_\lab$,
% the expression $r$ may be repeated an arbitrary number of times. 
The path $\pathvar$ is a set of label-to-int mappings that tracks
the current values assigned by Kleene closures during runtime
of either generation or validation.
On first entering a Kleene closure $r*_\lab$,
$\lab \rightarrow 0$ is entered into the path.
On each successive iteration, this mapping is incremented.

\mypara{Environment}
The environment $\env$ tracks how generation is assigning labels to values via bindings,
$(lab \Longleftarrow p(args, \lab)) \mapsto e$.
As the same label may be re-assigned different values in each iteration of a
(possibly nested) kleene closure,
the environment is really a mapping of (label, path) pairs to values.
We write $\env^\deps_\pathvar(\lab)$
to denote looking up the label \lab in \env,
given some path and dependency map
(because \deps is always the same for any given symbolic expression,
we omit it in future sections to reduce syntactic clutter.)
The way this works is that
we filter out all variables from
the path that are not in the \lab's dependencies.
This then gives us a $\pathvar'$ which contains only the
relevant context for the label,
and can be used for the lookup in the environment.

% \begin{example}
% Consider the expression:
% \[
% (n \Longleftarrow \mathit{True}) \mapsto ((x \Longleftarrow \mathit{True}) \mapsto x)*_n
% \]

% One list that could be generated from this would be $[8, 4, 17]$.
% The corresponding environment would be:
% \begin{align*}
%     \{  &(n, \{\}) \rightarrow 3, \\
%         &(x, \{ n \rightarrow 0 \}) \rightarrow 8, (x, \{ n \rightarrow 1 \}) \rightarrow 4, (x, \{ n \rightarrow 2 \}) \rightarrow 17 \}
% \end{align*}

% $n$ occurs outside of all kleene closures, and thus has an empty context.
% $x$ is chosen to be $8$, $4$, and $17$, respectively,
% on the $1^{st}$, $2^{nd}$ and $3^{rd}$ iterations of the unrolling of the kleene closure indexed by $n$,
% and thus has a separate value for each of these three contexts.
% \end{example}
% Since a single variable may take on multiple values in different contexts,
% the environment really maps labels to dictionaries which map contexts to values.

