\begin{enumerate}
    \item $\pathvar$ is our path, or ``context.'' It keeps track of which labels are taking on which values currently.
    
    \item $\env$ is our environment. It functions as a dictionary that maps labels to values. Since a single variable may take on multiple values in different contexts, the environment really maps labels to dictionaries which map contexts to values.
    
    \item $\deps$ is our dependency map. Since certain variable are only defined in the context of other variables, we need a map to tell us which other variable must be defined to bring another variable into scope. So the dependency maps labels to lists of labels for this purpose. For instance, consider the symbolic expression $a -> (b -> b)*_{a}$. Since $b$ is defined in the star expression labelled ``$a$'' then $b$ is only defined if $a$ is defined in the current context of an expression. Thus, $\envget{\deps}{b}=[a]$. When using the environment, we assume the dependency map is precomputed. We also assume the dependency map is consulted before looking up a variables specific value in the environment.
    
    \item $\envput{E}{\lab}{x}$ means ``Insert the key-value pair $(\lab, x)$ into $E$.'' if $E$ is $\env$ then we assume that this operation takes the current context and dependency map into account
    
    \item $\envget{E}{\lab}$ means ``get the value for key $x$ in $E$.'' If $E$ is $\env$ then we assume that this operation takes the current context and dependency map into account.
    
    \item $\envhas{E}{\lab}$ means ``check if $\lab$ is in the key set of $E$''
    
    \item $\hastype{x}{T}$ means ``$x$ is of type $T$'' where $T$ is either a base type $\base$, an arbitrary type $\complex$, or. A base type is a type that is not a list, an arbitrary type either a base type or a list of some arbitrary type. TODO, this isn't quite true I don't think... it really means ``surface expression $x$ will generate values of type $T$.'' maybe need new symbol for this?
    
    \item $e \gens x$ means ``symbolic expression $e$ generates value $x$''

    \item $x \translate y$ means ``surface expression $x$ translates to  core expression $y$''
    
    \item $xs ++ ys$ means ``append the list $ys$ to the list $xs$''
    
    \item $x :: xs$ means ``prepend the element $x$ to the list $xs$''
    
    \item $\fresh{x}$ means that value $x$ is a random value that will never be used again in any other expression
    
    \item $\arbitrary{x}$ means that value $x$ is a random value that may be used again in subsequent expressions.
\end{enumerate}
